---
title: "Peer Graded Assignment: Reproducible Research Course Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


## First read in the data

```{r, echo=TRUE}
library(ggplot2)
unzip("activity.zip")
activitydata<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
### Calculate the number of steps taken on each day
```{r,echo=TRUE}
TDaySteps<-aggregate(steps~date, data=activitydata, sum)
```

### Create histogram of the number of steps taken on each day
```{r,echo=TRUE}
hist(TDaySteps$steps, main=paste("Histogram of number of steps taken per day"), xlab="Steps")
```

### Calculate and report the mean and median of steps taken per day
``` {r,echo=TRUE}
TDayStepsMean<-mean(TDaySteps$steps)
print(TDayStepsMean)
TDayStepsMedian<-median(TDaySteps$steps)
print(TDayStepsMedian)
```

##What is the average daily activity pattern?

###Mean number of steps by day
```{r,echo=TRUE}
MIntervalSteps<-aggregate(steps~interval, data=activitydata, mean)
```
###Plot of average number of steps across days by interval
```{r,echo=TRUE}
plot(MIntervalSteps$interval, MIntervalSteps$steps, type="l", xlab="Interval", ylab="Steps", main="Average number of steps across days by interval")
```

###Identify the interval with the highest average number of steps per day
```{r,echo=TRUE}
print(MIntervalSteps[MIntervalSteps$steps == max(MIntervalSteps$steps), ] )
```
##Imputing missing values

###Calculate and report total number of missing values
```{r,echo=TRUE}
print(nrow(activitydata[is.na(activitydata$steps)|is.na(activitydata$date)|is.na(activitydata$interval), ] ))
```
###Impute data using the mean of that interval
```{r,echo=TRUE}
activitydata2<-activitydata
impute <- function(steps, interval) {
  complete <- NA
  if (!is.na(steps))
    complete <- c(steps)
  else
    complete <- (MIntervalSteps[MIntervalSteps$interval==interval, "steps"])
  return(complete)
}
activitydata2$steps <- mapply(impute, activitydata2$steps, activitydata2$interval)
```
###Calculate the number of steps taken on each day
```{r,echo=TRUE}
ImputedTDaySteps<-aggregate(steps~date, data=activitydata2, sum)
```
###Create histogram of the number of steps taken on each day
```{r,echo=TRUE}
hist(ImputedTDaySteps$steps, main=paste("Histogram of number of steps taken per day"), xlab="Steps")
```

###Calculate and report the mean and median of steps taken per day
```{r,echo=TRUE}
ImputedTDayStepsMean<-mean(ImputedTDaySteps$steps)
print(ImputedTDayStepsMean)
ImputedTDayStepsMedian<-median(ImputedTDaySteps$steps)
print(ImputedTDayStepsMedian)
```
##Are there differences in activity patterns between weekdays and weekends?

###Create weekday/weekend factor variable
```{r,echo=TRUE}
activitydata2$date<-strptime(activitydata2$date,"%Y-%m-%d")
activitydata2$weekday <- as.character(as.factor(weekdays(activitydata2$date)))
activitydata2$weekday[activitydata2$weekday=="Monday"] <- "Weekday"
activitydata2$weekday[activitydata2$weekday=="Tuesday"] <- "Weekday"
activitydata2$weekday[activitydata2$weekday=="Wednesday"] <- "Weekday"
activitydata2$weekday[activitydata2$weekday=="Thursday"] <- "Weekday"
activitydata2$weekday[activitydata2$weekday=="Friday"] <- "Weekday"
activitydata2$weekday[activitydata2$weekday=="Saturday"] <- "Weekend"
activitydata2$weekday[activitydata2$weekday=="Sunday"] <- "Weekend"
activitydata2$weekday<-as.factor(activitydata2$weekday)
```
##Plot of average number of steps across days by interval

###Mean number of steps by day
```{r,echo=TRUE}
ImputedMIntervalSteps<-aggregate(steps~interval*weekday, data=activitydata2, mean)
attach(ImputedMIntervalSteps)
par(mfrow=c(1,2))
p1 <- ggplot(ImputedMIntervalSteps, aes(x=interval, y=steps)) +
  geom_line(alpha=.3) +
  facet_grid(weekday ~ .) +
  ggtitle("Average number of steps across days by interval") +
  theme(legend.position="none") 
```
